---
title: "Exploratory Data Analysis"
author: "Clara Chua"
date: "9/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
packages = c('tidyverse','rgdal', 'maptools', 'sf','raster','spatstat', 'tmap', 'gridExtra', 'leaflet')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

<i>Work in progress</i>

## Introduction
This is the exploratory data analysis (EDA) for my capstone project, that looks into the geospatial analysis of airbnb rentals in the Singapore market.  

## Examining the Data

### What Data is Available?
Inside Airbnb provides information on the following:
* Listings - Summary information on listings
* Detailed Listings - Detailed listing information of airbnb for rent
* Calendar - Detailed calendar data for listings
* Reviews - Summary review data
* Detailed Reviews - Detailed review data for listings
* Neighbourhoods - list of neighbourhoods in the city and a neighbourhood GeoJSON file 

```{r, echo=TRUE, message=FALSE}
# Loading the Data
listings <- read_csv("../data/listings.csv") %>% as.data.frame()
d_listings <- read_csv("../data/detailedlistings.csv") %>% as.data.frame()
calendar <- read_csv("../data/calendar.csv") %>% as.data.frame()
reviews <- read_csv("../data/reviews.csv") %>% as.data.frame()
d_reviews <- read_csv("../data/detailedreviews.csv") %>% as.data.frame()
neighbourhoods <- read_csv("../data/neighbourhoods.csv") %>% as.data.frame()
```

```{r, echo=TRUE, message=FALSE}
# Reviewing the data
listings
d_listings
reviews
calendar
d_reviews
```

```{r, echo=TRUE, message=FALSE}
# Reviewing the data
filter(d_listings, require_guest_phone_verification == "TRUE")
n_distinct(reviews$listing_id)
length(reviews$listing_id)
min(reviews$date)
max(reviews$date)
```
#### Missing data
- Check for missing data, outliers (e.g. Review date < earliest listing date)
```{r, echo=TRUE, message=FALSE}
d_listings %>% summarise_all(funs(sum(is.na(.))))
d_reviews %>% summarise_all(funs(sum(is.na(.))))
listings %>% summarise_all(funs(sum(is.na(.))))
calendar %>% summarise_all(funs(sum(is.na(.))))
reviews %>% summarise_all(funs(sum(is.na(.))))
```
There are missing comments in the review column. 
There are also listings without reviews


```{r, echo=TRUE, message=FALSE} 
# Filter the missing data to see what to do with it
missing_list <- listings %>% 
                filter(is.na(listings$host_name))
missing_list

```



### Type of Accommodation available
The following plot shows the types of accommodation available and the average price of each accommodation type by region / neighbourhood.

#### By Region
Examining the listings by region, we see that unsurprisingly majority of the listings are in the Central Region, with more entire apartments for rent.  Interestingly, there are more listings of private rooms than entire apartments in the other regions.  We could surmise that these non-central region listings are possibly owner-occupied homes, who are renting out a spare room for additional income.  
```{r, echo=TRUE, message=FALSE}
# Plotting the type of accommodation and average price of accommodation by region / neighbourhood
regionlist <- listings %>%
                group_by(neighbourhood_group, room_type) %>%
                summarise(
                  num_listings = n(),
                  avg_price = mean(price),
                  med_price = median(price))
                # arrange(desc(avg_price))
regionlist
```
```{r, echo=TRUE, message=FALSE}
# Plotting the type of accommodation and average price of accommodation by region / neighbourhood
base_p <- ggplot(regionlist, aes(x=room_type, fill = room_type)) + theme(axis.text.x = element_blank())
p1 <- base_p + 
      geom_col(aes(y=num_listings)) +
      facet_grid(cols=vars(neighbourhood_group))
p2 <- base_p + geom_col(aes(y=med_price, color = room_type)) +
      facet_grid(cols=vars(neighbourhood_group))
# p2 <- ggplot(regionlist, aes(room_type, num_listings, fill = room_type)) +
#   geom_col() +
#   facet_grid(cols = vars(neighbourhood_group)) +
#   geom_jitter(aes(room_type, avg_price))
grid.arrange(p1, p2, ncol = 1)

```


#### Violin plot of prices
The violin plot shows that there are possibly outliers with bad data that we will need to exclude (e.g. $10,000 rental for an entire home/apt for a day)
```{r, echo=TRUE, message=FALSE}
# Do a violin plot of the prices of the various listings in each region?
ggplot(listings, aes(room_type, price, color=room_type))+ 
          geom_violin() +
          facet_grid(cols = vars(neighbourhood_group))+
          theme(axis.text.x = element_blank())
```


#### By Neighbourhood

```{r, echo=TRUE, message=FALSE}
# Plotting the type of accommodation and average price of accommodation by region / neighbourhood
nhoodlist <- listings %>%
                group_by(neighbourhood, room_type) %>%
                  summarise(
                    num_listings = n(),
                    avg_price = mean(price),
                    med_price = median(price)) %>%
                arrange(desc(med_price))
nhoodlist
```
The neighbourhood list shows some outlier where there is an entire home / apartment's rental in Tuas (not a central neighbourhood) is $10,001 per night.  As there is only 1 listing, we may wish to ignore this altogether.  We can also see that the neighbourhood with the most number of listings is Geylang and Kallang, which are adjacent areas.  This can be seen more clearly in the map of listings.

#### Read in neighbourhood maps / polygons
```{r, echo=TRUE, message=FALSE}
# reading in the neighbourhood geojson file
nhood_map <- readOGR(dsn = "../data/neighbourhoods.geojson", layer="neighbourhoods")
nhood_map2 <- st_read(dsn = "../data/neighbourhoods.geojson", layer="neighbourhoods")
nhood_map2
#plot(nhood_map)
```
#### Looking for outliers
The following code snippet is to clean data (currency to numeric) from the detailed listings and look for outliers
```{r, echo = TRUE, message = TRUE}
# Remove $ and , symbol in columns where currency is read as character.
strip_dollars = function(x) {as.numeric(gsub("[\\$,]", "", x))}
d_listings[,61:65] <- sapply(d_listings[,61:65], strip_dollars)
d_listings[,67] <- sapply(d_listings[,67], strip_dollars)
d_listings

#Checking out detailed listings with daily stay of > $2,000
outliers <- d_listings %>%
              filter(d_listings$price > 2000)
outliers

#Remove outliers from listings
nhoodlist_clean <- listings %>%
                    filter(listings$price < 2000) %>%
                    group_by(neighbourhood, room_type) %>%
                    summarise(
                      num_listings = n(),
                      avg_price = mean(price),
                      med_price = median(price)) %>%
                    arrange(desc(med_price))

```




### Hosts & Listings
This section for EDA on hosts
- Number of hosts and number of listings
- Time series of host join dates?
- Time series of listing dates (difference between yearly data)
```{r, echo=TRUE, message=FALSE}
list_byhost <- listings %>%
                group_by(host_id, host_name) %>%
                count(name = "number_of_listings", sort = TRUE) %>%
                #arrange(desc(num)) %>%
                ungroup()
  
list_byhost %>% group_by(number_of_listings) %>% count(name = "number_of_hosts")
```


### Reviews
This section for EDA on reviews
##### Inactive Listings (listings with no reviews)
``` {r}
# How many listings do not have reviews (i.e. no stays)
no_reviews <- filter(listings, is.na(last_review))
# listings[[is.na(listings$last_review)]]
nrow(no_reviews)
no_reviews
```
There are 2,835 listings without reviews.  We can look into them to see where they are located and the room type and other factors to see if there are things that could contribute to why they have no reviews / not been rented out.

#### Notes
New Hosts in 2020
There were 149 new hosts in 2020
- Additional work: To download archival information to extract the new listings added per year (or time period)
```{r, echo=TRUE, message=FALSE}
# Are these new listings?  Proxy - when did the hosts sign up?
no_reviews_host <- left_join(no_reviews, d_listings, by = "id")
nrow(filter(no_reviews_host, host_since < "2020-01-01"))
nrow(filter(no_reviews_host, host_since >= "2020-01-01"))
# max(no_reviews_host$host_since)
```

##### Additional review data to do
- Average/Median # reviews per listing
- Listings with highest number of reviews
- Look at review scores, number of reviews, host_verification, etc from detailed review data (correlation, etc)

#### Data preparation
- any additional preparation (e.g. removing outliers, etc)


## Spatial EDA


#### Visualising neighbourhoods on tmap
```{r, echo=TRUE, message=FALSE}
# Plotting neighbourhood listings on tmap
pal <- colorFactor(c("blue", "red", "green"), domain = c("Entire home/apt", "Shared room", "Private room"))

#tmap mode
tmap_mode("view")

listings_sf <- listings %>% 
                filter(listings$price < 2000) %>%
                st_as_sf(.,
  # st_as_sf(x = listings,
                        coords = c("longitude", "latitude"),
                        crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")

# Join neighbourhood mapping and dataframe
# Is this necessary?
listings_join <- left_join(nhood_map2, listings, by = c("neighbourhood"))

#Plotting points
tm_basemap(leaflet::providers$OpenStreetMap) +
tm_shape(nhood_map2) +
  tm_polygons(alpha = 0.2) +
  tm_shape(listings_sf) +
  tm_dots(col = "room_type")
  # tm_dots() +
  # tm_shape(nhood_map2)
```
The listings are clearly clustered around the central regin of Singapore, close to the CBD and main shopping districts.  There are also listings in suburban parts of Singapore which, while less expected could be explained by either being closer to 'desirable' neighbourhoods such as East Coast, or closer to specific industrial areas such as Pioneer, Sembawang, Changi Business Park / Loyang.    

Check out the things in the cachement areas - should not be correct.  

#### Mapping median price, average price and number of listings
``` {r, echo= TRUE, message = FALSE}
# Join neighbourhood mapping and num listings and price dataframe
listings_join2 <- left_join(nhood_map2, nhoodlist_clean, by = c("neighbourhood"))

tmap_mode("view")
# REDO THIS ONE
# Show med price, ave price and num listings in a map layout
base <- tm_basemap(leaflet::providers$OpenStreetMap)
map1 <- base + tm_shape(listings_join2) + 
  tm_polygons("med_price")
# map2 <- base + tm_shape(listings_join2) + 
  # tm_polygons("avg_price", alpha = 0.2)
map3 <- base + tm_shape(listings_join2) + 
  tm_polygons("num_listings", palette = "Purples", alpha = 0.4)
tmap_arrange(map1, map3)
  # tm_fill("num_listings") + 
  # tm_fill("med_price")
```
From the maps above, we can also see that there are neighbourhoods without any data.  If we examine these closer, we can see that these are purely industrial zoned areas, without any residential zones (e.g. Jurong Island, Tuas, Western Islands, Lim Chu Kang, Seletar, Changi).  The neighbourhood designated as "Paya Lebar" encloses the airbase, and not what most people would recongise as Paya Lebar.  

### Spatial Point Pattern Analysis (SPPA)


#### Data Wrangling for SPPA
The following code translate the data that we have into the formats required for SPPA using the spatstat package.  The listings dataframe needs to be converted from lat-long to SVY21 format to be able to put the coordinates together.  It then needs to be converted to a SpatialPoints object for use in Spatstat.  
The polygon layer also needs to have the right projection (3414) and be converted to a SpatialPolygons object.  

Once done, we can then convert them into ppp and owin formats for spatstat to compute the density.  
``` {r, echo = TRUE, message = FALSE}
# listings to be converted from the listings_sf dataframe, to SVY21 format (otherwise it will be out of bounds when using Spatstat) and converting to SpatialPoints (sp)
# Polygon layer was previously imported as nhood_map, to add a projection and convert to SpatialPolygons 

listings_sf <- st_transform(listings_sf, 3414)
listings_sp <- st_coordinates(listings_sf) %>% #as.data.frame() %>%
                #dplyr::rename(longitude = X, latitude = Y) %>% 
                SpatialPoints(., proj4string = CRS("+init=epsg:3414")) 
nhood_sp <- spTransform(nhood_map, CRS("+init=epsg:3414")) %>%
            as(., "SpatialPolygons")
summary(listings_sp)
summary(nhood_sp)
plot(listings_sp)
plot(nhood_sp)

# Converting sp format to ppp format to be used in spatstat
listings_ppp <- as.ppp(listings_sp)
nhood_owin <- as(nhood_sp, "owin")
plot(nhood_owin)
summary(listings_ppp)
```
Here we create jittering and combine the listings and neighbourhoods to feed into point pattern analysis.  

``` {r echo = TRUE, eval = TRUE}
# create jittering
listing_ppp_jit <- rjitter(listings_ppp, retry = TRUE, nsim = 1, drop = TRUE)
plot(listing_ppp_jit)
summary(listing_ppp_jit)
summary(nhood_owin)


# Using subzone mapping instead to test (will remove later)

# subzone <- readOGR(dsn = "../data", layer="MP14_SUBZONE_WEB_PL")
# crs(subzone)
# sg_sp <- as(subzone, "SpatialPolygons")
# plot(sg_sp)
# summary(sg_sp)
# sg_owin <- as(sg_sp, "owin")
# plot(sg_owin)
# summary(sg_owin)
# listingsSG_ppp = listing_ppp_jit[sg_owin]
# plot(listingsSG_ppp)
# summary(listingsSG_ppp)


# Combining listings and the neighbourhoods
listingsg_ppp = listing_ppp_jit[nhood_owin]
plot(listingsg_ppp)
# summary(listingsg_ppp)
# listingsg_ppp.km <- rescale(listingsg_ppp, 1000, "km")
# kde_listings_bw <- density(listingsg_ppp.km)

```

#### Quadrat Analysis

For the overall listings for SG, we can run the quadrat test to see if there 


```{r, echo=TRUE, message=FALSE}
qtest <- quadrat.test(listingsg_ppp, 
                   nx = 20, ny = 15)
qtest

plot(listingsg_ppp)
plot(qtest, add = TRUE, cex =.1)
```

Conditional Monte Carlo test of CSR using quadrat counts
```{r, echo=TRUE, message=FALSE}
qtest_MC <- quadrat.test(listingsg_ppp, 
                   nx = 20, ny = 15,
                   method="M",
                   nsim = 999)
qtest_MC
```

#### Kernel Density Estimation
Plotting the KDE plots
```{r, echo=TRUE, message=FALSE}
# Rescaling ppp to m instead of km and plotting the kernel density estimation
listingsg_ppp.km <- rescale(listingsg_ppp, 1000, "km")
kde_listings_bw <- density(listingsg_ppp, sigma=bw.diggle, edge = TRUE, kernel = "gaussian")
plot(kde_listings_bw)
kde_listings_bw_km <- density(listingsg_ppp.km, sigma=bw.diggle, edge = TRUE, kernel="gaussian")
plot(kde_listings_bw_km)
```
As we can see from above, a higher density of listings are within the city centre and we can take a closer look at those areas.

#### Using Adaptive Bandwidth
```{r, echo=TRUE, message=FALSE}
kde_listings_adaptive <- adaptive.density(listingsg_ppp, method = "kernel")
kde_listings_adaptive2 <- adaptive.density(listingsg_ppp.km, method = "kernel")
plot(kde_listings_adaptive2)
```

#### Mapping KDE onto map to identify areas for further investigation
``` {r, echo=TRUE, message = FALSE}
gridded_kde_listings_bw <- as.SpatialGridDataFrame.im(kde_listings_bw)
gridded_kde_listings_adapt <- as.SpatialGridDataFrame.im(kde_listings_adaptive)
spplot(gridded_kde_listings_bw)
spplot(gridded_kde_listings_adapt)
kde_listings_bw_raster <- raster(gridded_kde_listings_bw)
kde_listings_adap_raster <- raster(gridded_kde_listings_adapt)
projection(kde_listings_bw_raster) <- CRS("+init=EPSG:3414")
projection(kde_listings_adap_raster) <- CRS("+init=EPSG:3414")
kde_listings_bw_raster
kde_listings_adap_raster
```


```{r, echo=TRUE, message=FALSE}
tm_basemap(leaflet::providers$OpenStreetMap) +
  tm_shape(nhood_map2) +
  tm_polygons(alpha = 0.1) +
tm_shape(kde_listings_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```
From the raster map (using the bw.diggle function), we can see that the following neighbourhoods have higher than normal density
Novena, Kallang, Geylang, Rochor, River Valley, Outram / Bukit Merah (fringe), Downtown Core

```{r echo=TRUE, eval=TRUE}
# tm_basemap(leaflet::providers$OpenStreetMap) +
  tm_shape(nhood_map2) +
  tm_polygons(alpha = 0.1) +
tm_shape(kde_listings_adap_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```
The adaptive raster gives a few more areas to investigate as well: Orchard, Singapore River.


#### Clark and Evans Test for all listings
```{r, echo=TRUE, message=FALSE}
# 999 simulations take too long.  Only do CE test on smaller polygons
# clarkevans.test(listingsg_ppp.km,
#                 correction="none",
#                 clipregion=NULL,
#                 alternative=c("two.sided"),
#                 nsim=99)
```

##### Using different fixed kernel sizes (do we need to?) 
Density mapping using different fixed kernel sizes or 500m, and 750m.  

```{r echo=TRUE, eval=TRUE}
# kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
# plot(kde_childcareSG_600)

kde_listings_500 <- density(listingsg_ppp.km, sigma=0.5, edge=TRUE, kernel="gaussian") 
plot(kde_listings_500)

kde_listings_750 <- density(listingsg_ppp.km, sigma=0.75, edge=TRUE, kernel="gaussian") 
plot(kde_listings_750)
```

##### Notes from meeting 21/10/2020
What is the advantage of 1 over the other one (adaptive kernel density may be more useful than the fixed kernel density) - both also have its reasons to use by using fixed kernel - help us to get some good insights also. Depends on what we're trying to catch.  

Covariate - slightly different.  Cross function - to see that if we look at 3 types of listing (room type) - cross functions, those listed whole unit same as those who listed the partial unit.  
Point events - listing of whole unit - co-exist together or located at different locations

Another way to look at the cross - whether this listing - are they near to existing hotels or hostels (formal accommodation)

Point process - one with locations of all hotels vs listings - do they locate together or not.
Cross function - do they co-exist.  Look at all airbnb listing - but look at subset of them.  Airbnb listing - vs existing hotel or location of tourist attraction or proximity to public transport (MRT stations) - investigate the patterns.  

Logical ones - for tourists - taking bus is v unlikely but taking MRT will be easy.  Proximity to large shopping malls, proximity to tourist attractions.  

Don't exclude the other listings. 
Do kde for different room types, and look at distribution of rental prices - lower 25, middle 50 and top 25 percent, can see different spatial point patterns - can be viewed as kernel density map.  Where are the less expensive listings distributed around, and where are the more expensive types.  

Once we cut out 2nd order - still apply kernel density to show the localised density maps.  When you do the 2nd order, can more specifically describe where they tend to cluster to compute the KDE map in the small area.  

#### Subsetting into all polygons (except polygons with missing values)
This section is for writing the function to subset all polygons and perform KDE for polygons with listings
```{r, echo=TRUE, message=FALSE}

```


```{r, echo=TRUE, message=FALSE}
# code
```

```{r, echo=TRUE, message=FALSE}
# code
```

```{r, echo=TRUE, message=FALSE}
# code
```
